If you see something, say something (to others): A spread of misinformation and socially-contingent corrections within social networks

### 2018-02-06 manuscript_1.0.tex
Update on Overleaf.

\note{%Word count: \quickwordcount \\
      \textbf{Draft in progress. Please do not cite without permission.} Please direct any questions and inquiries to \href{mailto:hyunjin.song@univie.ac.at}{hyunjin.song@univie.ac.at}}

@@

Against this backdrop, our goal in this contribution is to offer a more systematic assessment of underlying mechanisms of misinformation spreading and its correction, focusing on one's \emph{social} contexts in which such exposure to (mis)information and corrective messages are received and processed. We argue while \emph{exposure} to (mis)information is likely to follow a simple contagion process, \emph{changes} in one's beliefs regarding such (mis)information -- which ultimately \emph{the} goal of corrective messages -- is likely to be, in \citeauthor{centola2007complex}'s (\citeyear{centola2007complex}) term, a \enquote{complex contagion} where such changes require multiple sources of affirmation and reinforcement compared to simple contagion process. As a result, the effects of fact-checking and corrective messages are likely to be highly \emph{socially} contingent, and we argue that vast of prior studies have not considered this possibility to date \parencite[for a notable exception, see][]{margolin2017}.

In what follows, we first review existing evidence regarding political misperceptions and the effect of fact-checking messages. We advance our dynamic system model perspective by combining an individual-level cognitive and affective basis of adopting such misinformation with a social context of which an exposure to misinformation and corrections are received. Based on a well-known class of an epidemic model of virus infection and recovery, we propose an integrated model of misinformation diffusion and socially-contingent corrections on social networks, with a special focus on the differences between simple contagion of misinformation and complex contagion of corrections. Relying on Agent-based simulations, we further explore various boundary conditions of such dynamics, aiming to uncover how and when such misinformation propagates into the public, as well as what factors facilitate or hinder such diffusion process.

>>>>

\note{Word count: \quickwordcount \\
      \textbf{Draft in progress. Please do not cite without permission.} Please direct any questions and inquiries to \href{mailto:hyunjin.song@univie.ac.at}{hyunjin.song@univie.ac.at}}

@@

Against this backdrop, our goal in this contribution is to offer a more systematic assessment of underlying mechanisms of misinformation spreading and its correction, focusing on one's \emph{social} contexts in which such exposure to (mis)information and corrective messages are received and processed. We argue while \emph{exposure} to (mis)information is likely to follow a simple contagion process, \emph{changes} in one's beliefs regarding such (mis)information -- which ultimately \emph{the} goal of corrective messages -- is likely to be, in \citeauthor{centola2007complex}'s (\citeyear{centola2007complex}) term, a \enquote{complex contagion} where such changes require multiple sources of affirmation and reinforcement compared to simple contagion process. As a result, the effects of fact-checking and corrective messages are likely to be highly \emph{socially} contingent, and we argue that vast of prior studies have not considered this possibility to date \parencite[yet for a notable exception, see][]{margolin2017}.

In what follows, we first review existing evidence regarding political misperceptions and the effect of fact-checking messages. We advance our dynamic system perspective by combining an individual-level cognitive and affective basis of adopting such misinformation with a social context of which an exposure to misinformation and corrections are received. Based on a well-known class of an epidemic model of virus infection and recovery, we propose an integrated model of misinformation diffusion and socially-contingent corrections on social networks, with a special focus on the differences between simple contagion of misinformation and complex contagion of corrections. Relying on Agent-based simulations, we further explore boundary conditions of such dynamics, aiming to uncover how and when such misinformation propagates into the public, as well as what factors facilitate or hinder such diffusion process.
